---
output: 
  github_document:
    pandoc_args: --webtex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(refund)
library(reshape2)
library(tidyfun)
library(ggplot2)
source("./function/quadWeights.R")
source("./function/fpca.tfd.R")
```

## FDA Functions for tf-Class

## Introduction

**`tidyfun`** is geared toward making functional data analysis (FDA) easy, especially for data wrangling and exploratory analysis. 

Look [here](https://fabian-s.github.io/tidyfun/) for an introduction on **`tidyfun`**  with examples.

Further examples on data wrangling and plotting with tdf class on this github page can be found in the data preparation folder 
[here](https://github.com/gekim0519/tidyfun_fpca/blob/master/data_preparation/preparation.md).

## Objective

- The role of this project is to extend functions for FDA. Specially, four functions were made for conducting fpca and function-on-scalar regression on tfd data types.
- The functions are modified from the **`refund`** package ([link]( https://github.com/refunders/refundable)) to take in a dataframe containing tfd columns directly as an input.

## Data Description

The type of data we will be working with is **`tf`-Class** data and more specifically, **`tfd`** objects. **`tfd`** is one of the two subclasses of **`tf`-Class**, a new data type that allows functional data to be stored as vectors. More information on **`tf`-Class** can be found [here](https://fabian-s.github.io/tidyfun/articles/x01_Intro.html).

**Example Dataset - DTI, Diffusion Tensor Imaging**

- Fractional anisotropy (FA) tract profiles for the corpus callosum (cca) and the right corticospinal tract (rcst).
- The MRI/DTI data were collected at Johns Hopkins University and the Kennedy-Krieger Institute.
- `DTI` has 382 rows and 9 columns including two functional covariates (`cca`, `rcst`) in forms of matrices. 
- Using tidyfun, new dataframe `dti` has cca and rcst in tfd type.
- Link to further description: [rdrr.io/cran/refund/man/DTI.html](rdrr.io/cran/refund/man/DTI.html)

```{r}
DTI = refund::DTI

dti = with(refund::DTI, 
  data.frame(id = ID, sex = sex, 
    case = factor(ifelse(case, "MS", "control")))) %>% as.tbl %>% 
        mutate(cca = tfd(DTI$cca, seq(0,1, l = 93), signif = 2) %>%
                     tfd(arg = seq(0,1,l = 93)),
               rcst = tfd(DTI$rcst, seq(0, 1, l = 55), signif = 3))

dti %>%
  head()
```

Let's view cca in a spaghetti plot.

```{r}
dti %>%  
  ggplot(aes(y = cca, color = factor(id))) + 
  geom_spaghetti(alpha = .3) + xlab("days") + theme(legend.position="none") + ggtitle("FA tract profiles from the corpus callosum")
```

## Functions and Usage

#### Functional principal components analysis (FPCA) by smoothed covariance

**Model**

$$ 
Y_i(t) = \mu(t) + \sum C_{ik} \phi_{k}(t) + \epsilon_i(t) 
$$

**`fpca.tdf`**

- Decomposes functional observations using functional principal components analysis. A mixed model framework is used to estimate scores and obtain variance estimates.
- Altered from `refund::fpca.sc`, fpca.tfd  allows users to apply the function directly on dataframes with a tfd column, specified in the `col` argument.
- Uses penalized splines to smooth the covariance function, as developed by Di et al. (2009) and Goldsmith et al. (2013).

#### Example

```{r}
# DTI
fit.cca = fpca.tfd(data = dti, col = cca)

fit.mu = data.frame(mu = fit.cca$mu,
                    n = 1:ncol(fit.cca$Yhat))
fit.basis = data.frame(phi = fit.cca$efunctions, #the FPC basis functions.
                       n = 1:ncol(fit.cca$Yhat))
```

Let's plot the estimated mean function of `cca`.

```{r}
## plot estimated mean function
ggplot(fit.mu, aes(x = n, y = mu)) + geom_path() + theme_bw() + ggtitle("Estimated mean function of corpus callosum")
```

There are nine basis eigenfunctions in `fit.basis`. Let's plot the first two.

```{r}
## plot the first two estimated basis functions
fit.basis.m = melt(fit.basis, id = 'n')
ggplot(subset(fit.basis.m, variable %in% c('phi.1', 'phi.2')), aes(x = n,
y = value, group = variable, color = variable)) + geom_path() + theme_bw() + ggtitle("First two estimated FPC basis functions")
```

#### Function-on-Scalar Regression (FoSR) 

**Model**

$$
y_i(t) = \beta_{0}(t) + \sum x_{ik} \beta_{k}(t) + \epsilon_i(t)
$$

**ols_cs_tfd**

- Fitting function for FoSR for cross-sectional data, estimates model parameters using GLS.
- We edited `refund::ols_cs`. While the inputs did not change, `ols_cs_tfd`’s argument, data will be a dataframe with `tfd` type column, which will be the response of the proposed model.



`



