---
title: "explore tidyfun"
author: "Gaeun Kim"
date: "1/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidyfun)
library(refund)
library(ggplot2)
library(reshape2)
library(mgcv)
source("./function/quadWeights.R")
source("./function/fpca.tfd.R")
```

read in datasets
```{r}
temp_tfd = load("data/temp_tfd.RData")
handw_tfd= load("data/handw_tfd.RData")

DTI = refund::DTI

dti = with(refund::DTI, 
  data.frame(id = ID, sex = sex, 
    case = factor(ifelse(case, "MS", "control")))) %>% as.tbl %>% 
        mutate(cca = tfd(DTI$cca, seq(0,1, l = 93), signif = 2) %>%
                     tfd(arg = seq(0,1,l = 93)),
               rcst = tfd(DTI$rcst, seq(0, 1, l = 55), signif = 3))
```

fpca.tfd function examples
```{r}
fit.cca = fpca.tfd(data = dti, col = rcst)

fit.mu = data.frame(mu = fit.cca$mu,
                    n = 1:ncol(fit.cca$Yhat))
fit.basis = data.frame(phi = fit.cca$efunctions, #the FPC basis functions.
                       n = 1:ncol(fit.cca$Yhat))

## plot estimated mean function
ggplot(fit.mu, aes(x = n, y = mu)) + geom_path() 

## plot the first two estimated basis functions
fit.basis.m = melt(fit.basis, id = 'n')
ggplot(subset(fit.basis.m, variable %in% c('phi.1', 'phi.2')), aes(x = n,
y = value, group = variable, color = variable)) + geom_path()

```

putting cd4 dataset to fpca.tfd function

```{r}
data(cd4)
# CD4 cell counts for 366 subjects between months -18 and 42 since seroconversion. Each subject's observations are contained in a single row.
# subject * weeks

cd4_tfd = cd4 %>%
  tfd()
cd4_df = data_frame(cd4_tfd, id = 1:366) %>%
  select(cd4 = cd4_tfd, everything())

fit.cd4 = cd4_df %>%
  fpca.tfd(data =., col = cd4)

fit.mu = data.frame(mu = fit.cd4$mu,
                    n = 1:ncol(fit.cd4$Yhat))
fit.basis = data.frame(phi = fit.cd4$efunctions, #the FPC basis functions.
                       n = 1:ncol(fit.cd4$Yhat))

## plot estimated mean function
ggplot(fit.mu, aes(x = n, y = mu)) + geom_path() 
```

cd4 example on the refund page
```{r}
Fit.MM = fpca.sc(cd4, var = TRUE, simul = TRUE)

Fit.mu = data.frame(mu = Fit.MM$mu,
                    d = as.numeric(colnames(cd4)))
Fit.basis = data.frame(phi = Fit.MM$efunctions, # d × npc matrix of estimated eigenfunctions of the functional covariance, i.e., the FPC basis functions.
                       d = as.numeric(colnames(cd4)))

## for one subject, examine curve estimate, pointwise and simultaneous itervals
EX = 1
EX.MM = data.frame(fitted = Fit.MM$Yhat[EX,],
           ptwise.UB = Fit.MM$Yhat[EX,] + 1.96 * sqrt(Fit.MM$diag.var[EX,]),
           ptwise.LB = Fit.MM$Yhat[EX,] - 1.96 * sqrt(Fit.MM$diag.var[EX,]),
           simul.UB = Fit.MM$Yhat[EX,] + Fit.MM$crit.val[EX] * sqrt(Fit.MM$diag.var[EX,]),
           simul.LB = Fit.MM$Yhat[EX,] - Fit.MM$crit.val[EX] * sqrt(Fit.MM$diag.var[EX,]),
           d = as.numeric(colnames(cd4)))

## plot data for one subject, with curve and interval estimates
EX.MM.m = melt(EX.MM, id = 'd')
ggplot(EX.MM.m, aes(x = d, y = value, group = variable, color = variable, linetype = variable)) +
  geom_path() +
  scale_linetype_manual(values = c(fitted = 1, ptwise.UB = 2,
                        ptwise.LB = 2, simul.UB = 3, simul.LB = 3)) +
  scale_color_manual(values = c(fitted = 1, ptwise.UB = 2,
                     ptwise.LB = 2, simul.UB = 3, simul.LB = 3)) +
  labs(x = 'Months since seroconversion', y = 'Total CD4 Cell Count')

## plot estimated mean function
ggplot(Fit.mu, aes(x = d, y = mu)) + geom_path() +
  labs(x = 'Months since seroconversion', y = 'Total CD4 Cell Count')

## plot the first two estimated basis functions
Fit.basis.m = melt(Fit.basis, id = 'd')
ggplot(subset(Fit.basis.m, variable %in% c('phi.1', 'phi.2')), aes(x = d,
y = value, group = variable, color = variable)) + geom_path()

```
looks same as the tfd version!

bayes_fosr:
"Wrapper function that implements several approaches to Bayesian function on scalar regression."

Exammple code from refund

* I am getting errors regarding gibbs so I removed it from the sample for now.
* error message: Error in tx[, ind.cur] %*% siginv : non-conformable arguments
```{r}
##### Cross-sectional real-data examples #####

## organize data
#data(DTI)
DTI = subset(DTI, select = c(cca, case, pasat))
DTI = DTI[complete.cases(DTI),]
DTI$gender = factor(sample(c("male","female"), dim(DTI)[1], replace = TRUE))
DTI$status = factor(sample(c("RRMS", "SPMS", "PPMS"), dim(DTI)[1], replace = TRUE))

## fit models
default = bayes_fosr(cca ~ pasat, data = DTI)
VB = bayes_fosr(cca ~ pasat, data = DTI, Kp = 4, Kt = 10)
#Gibbs = bayes_fosr(cca ~ pasat, data = DTI, Kt = 10, est.method = "Gibbs", cov.method = #"Wishart", N.iter = 500, N.burn = 200)
OLS = bayes_fosr(cca ~ pasat, data = DTI, Kt = 10, est.method = "OLS")
GLS = bayes_fosr(cca ~ pasat, data = DTI, Kt = 10, est.method = "GLS")

## plot results
models = c("default", "VB", "OLS", "GLS")
intercepts = sapply(models, function(u) get(u)$beta.hat[1,])
slopes = sapply(models, function(u) get(u)$beta.hat[2,])

plot.dat = melt(intercepts); colnames(plot.dat) = c("grid", "method", "value")
ggplot(plot.dat, aes(x = grid, y = value, group = method, color = method)) + 
   geom_path() + theme_bw()

plot.dat = melt(slopes); colnames(plot.dat) = c("grid", "method", "value")
ggplot(plot.dat, aes(x = grid, y = value, group = method, color = method)) + 
   geom_path() + theme_bw()

#Gibbs - Error in tx[, ind.cur] %*% siginv : non-conformable arguments
# removed Gibbs
```

ols_cs:
"Fitting function for function-on-scalar regression for cross-sectional data."
```{r}
dti.ols = ols_cs(cca ~ pasat, data = DTI, Kt = 10)
```

```{r}
library(splines)
library(pbs)
```

```{r}
ols_cs_prac = function(formula, data=NULL, Kt=5, basis = "bs", verbose = TRUE){
  
  call <- match.call()
  tf <- terms.formula(formula, specials = "re")
  trmstrings <- attr(tf, "term.labels")
  specials <- attr(tf, "specials")
  where.re <-specials$re - 1
  if (length(where.re) != 0) {
    mf_fixed <- model.frame(tf[-where.re], data = data)
    formula = tf[-where.re]
    responsename <- attr(tf, "variables")[2][[1]]
    ###
    REs = list(NA, NA)
    REs[[1]] = names(eval(parse(text=attr(tf[where.re], "term.labels")), envir=data)$data)
    REs[[2]]=paste0("(1|",REs[[1]],")")
    ###
    formula2 <- paste(responsename, "~", REs[[1]], sep = "")
    newfrml <- paste(responsename, "~", REs[[2]], sep = "")
    newtrmstrings <- attr(tf[-where.re], "term.labels")
    formula2 <- formula(paste(c(formula2, newtrmstrings), 
                              collapse = "+"))
    newfrml <- formula(paste(c(newfrml, newtrmstrings), collapse = "+"))
    mf <- model.frame(formula2, data = data)
    if (length(data) == 0) {
      Z = lme4::mkReTrms(lme4::findbars(newfrml), fr = mf)$Zt
    }
    else {
      Z = lme4::mkReTrms(lme4::findbars(newfrml), fr = data)$Zt
    }
  }
  else {
    mf_fixed <- model.frame(tf, data = data)
  }
  mt_fixed <- attr(mf_fixed, "terms")
  
  # get response (Y)
  Y <- model.response(mf_fixed, "numeric")
  
  # x is a matrix of fixed effects
  # automatically adds in intercept
  X <- model.matrix(mt_fixed, mf_fixed, contrasts)
  
  ### model organization ###
  D = dim(Y)[2]
  I = dim(X)[1]
  p = dim(X)[2]
  
  if(basis == "bs"){
    Theta = bs(1:D, df = Kt, intercept=TRUE, degree=3)
  } else if(basis == "pbs"){
    Theta = pbs(1:D, df = Kt, intercept=TRUE, degree=3)
  }

  X.des = X
  Y.vec = as.vector(t(Y)) 
  X = kronecker(X.des, Theta)
  n.coef = dim(X.des)[2]
  
  ## OLS model fitting and processing results
  if(verbose) { cat("Using OLS to estimate model parameters \n") }
  model.ols = lm(Y.vec ~ -1 + X)
  Bx.ols = matrix(model.ols$coef, nrow = Kt, ncol = n.coef)  
  beta.hat.ols = t(Bx.ols) %*% t(Theta)
    
  resid.mat = matrix(resid(model.ols), I, D, byrow = TRUE)
    
  ## Get Residual Structure using FPCA
  ## note: this is commented out because, in simulations based on the headstart data, 
  ## using FPCA lead to higher-than-nominal sizes for tests of nested models. 
  ## using the raw covariance worked better. using FPCA is possible, but relies
  ## on some case-specific choices.
  # raw.resid.cov = cov(resid.mat)
  # fpca.resid = fpca.sc(resid.mat, pve = .9995, nbasis = 20)
  # resid.cov = with(fpca.resid, efunctions %*% diag(evalues) %*% t(efunctions))
    
  ## account for (possibly non-constant) ME nugget effect
  # sm.diag = Theta %*% solve(crossprod(Theta)) %*% t(Theta) %*% (diag(raw.resid.cov) - diag(resid.cov))
  # if(sum( sm.diag < 0 ) >0) { sm.diag[ sm.diag < 0] = min((diag(raw.resid.cov) - diag(resid.cov))[ sm.diag < 0])}
  # diag(resid.cov) = diag(resid.cov) + sm.diag
    
  sigma = cov(resid.mat) * (I - 1) / (I - p)
  
  ## get confidence intervals
  beta.UB = beta.LB = matrix(NA, p, D)
  for(p.cur in 1:p){
    ## confidence intervals for this model shouldn't be trusted
  }
  
  Yhat = X.des %*% beta.hat.ols
  
  ret = list(beta.hat.ols, beta.UB, beta.LB, Yhat, mt_fixed, data, model.ols, sigma)
  names(ret) = c("beta.hat", "beta.UB", "beta.LB", "Yhat", "terms", "data", "model.ols", "sigma")
  class(ret) = "fosr"
  ret
    
}
```

```{r}
dti.ols.prac = ols_cs_prac(cca ~ pasat, data = DTI, Kt = 10)
```


gibbs_cs_fpca:
"Cross-sectional FoSR using a Gibbs sampler and FPCA"
```{r}
gibbs_cs_fpca(cca ~ pasat, data = DTI, Kt = 10)
gibbs_cs_fpca_prac(cca ~ pasat, data = DTI, Kt = 10)

#keeps on running...
```

```{r}
library(MASS)
library(lme4)
gibbs_cs_fpca_prac = function(formula, Kt=5, Kp=2, data=NULL, verbose = TRUE, N.iter = 5000, N.burn = 1000, SEED = NULL, 
                         sig2.me = .01, alpha = .1, Aw = NULL, Bw = NULL, Apsi = NULL, Bpsi = NULL){

  call <- match.call()
  tf <- terms.formula(formula, specials = "re")
  trmstrings <- attr(tf, "term.labels")
  specials <- attr(tf, "specials")
  where.re <-specials$re - 1
  if (length(where.re) != 0) {
    mf_fixed <- model.frame(tf[-where.re], data = data)
    formula = tf[-where.re]
    responsename <- attr(tf, "variables")[2][[1]]
    ###
    REs = list(NA, NA)
    REs[[1]] = names(eval(parse(text=attr(tf[where.re], "term.labels")), envir=data)$data)
    REs[[2]]=paste0("(1|",REs[[1]],")")
    ###
    formula2 <- paste(responsename, "~", REs[[1]], sep = "")
    newfrml <- paste(responsename, "~", REs[[2]], sep = "")
    newtrmstrings <- attr(tf[-where.re], "term.labels")
    formula2 <- formula(paste(c(formula2, newtrmstrings), 
                              collapse = "+"))
    newfrml <- formula(paste(c(newfrml, newtrmstrings), collapse = "+"))
    mf <- model.frame(formula2, data = data)
    if (length(data) == 0) {
      Z = lme4::mkReTrms(lme4::findbars(newfrml), fr = mf)$Zt
    }
    else {
      Z = lme4::mkReTrms(lme4::findbars(newfrml), fr = data)$Zt
    }
  }
  else {
    mf_fixed <- model.frame(tf, data = data)
  }
  mt_fixed <- attr(mf_fixed, "terms")
  
  # get response (Y)
  Y <- model.response(mf_fixed, "numeric")
  
  if(!is.null(SEED)) { set.seed(SEED) }
  
  # x is a matrix of fixed effects
  # automatically adds in intercept
  W.des = X <- model.matrix(mt_fixed, mf_fixed, contrasts)
  
  ## subject covariates
  I = dim(X)[1]
  D = dim(Y)[2]
  p = dim(X)[2]
  
  ## bspline basis and penalty matrix
  Theta = bs(1:D, df=Kt, intercept=TRUE, degree=3)
  
  diff0 = diag(1, D, D)
  diff2 = matrix(rep(c(1,-2,1, rep(0, D-2)), D-2)[1:((D-2)*D)], D-2, D, byrow = TRUE)
  P0 = t(Theta) %*% t(diff0) %*% diff0 %*% Theta
  P2 = t(Theta) %*% t(diff2) %*% diff2 %*% Theta
  P.mat = alpha * P0 + (1-alpha) * P2
  
  ## data organization; these computations only need to be done once   
  Y.vec = as.vector(t(Y))
  t.designmat.W = t(kronecker(W.des, Theta))
  sig.W = kronecker(t(W.des) %*% W.des, t(Theta)%*% Theta)
  
  ## initial estimation and hyperparameter choice
  vec.BW = solve(kronecker(t(W.des)%*% W.des, t(Theta) %*% Theta)) %*% t(kronecker(W.des, Theta)) %*% Y.vec
  mu.q.BW = matrix(vec.BW, Kt, p)
  
  Yhat = as.matrix(W.des %*% t(mu.q.BW) %*% t(Theta))
  
  Aw = ifelse(is.null(Aw), Kt/2, Aw)
  if(is.null(Bw)){
    Bw = b.q.lambda.BW = sapply(1:p, function(u) max(1, .5*sum(diag( t(mu.q.BW[,u]) %*% P.mat %*% (mu.q.BW[,u])))))
  } else {
    Bw = b.q.lambda.BW = rep(Bw, p)
  }
  
  Apsi = ifelse(is.null(Apsi), Kt/2, Apsi)
  Bpsi = ifelse(is.null(Bpsi), Kt/2, Bpsi)
  Asig = 1; Bsig = 1
  
  ## matrices to store within-iteration estimates 
  BW = array(NA, c(Kt, p, N.iter))
      BW[,,1] = bw = matrix(0, Kt, p)
  BPSI = array(NA, c(Kt, Kp, N.iter))
      BPSI[,,1] = bpsi = matrix(0, Kt, Kp)
  C = array(NA, c(I, Kp, N.iter))
      C[,,1] = c.mat = matrix(rnorm(I*Kp, 0, .01), I, Kp)
  SIGMA = rep(NA, N.iter)
      SIGMA[1] = sig2.me = sig2.me
  LAMBDA.BW = matrix(NA, nrow = N.iter, ncol = p)
      LAMBDA.BW[1,] = lambda.bw = rep(1, p)
  LAMBDA.PSI = matrix(NA, nrow = N.iter, ncol = Kp)
      LAMBDA.PSI[1,] = lambda.psi = rep(1,Kp)
  y.post = array(NA, dim = c(I, D, (N.iter - N.burn))) 


  ## initialize estimates of fixed, random and pca effects
  beta.cur = t(bw) %*% t(Theta)
  fixef.cur = W.des %*% beta.cur
  psi.cur = t(bpsi) %*% t(Theta)
  pcaef.cur = c.mat %*% psi.cur

  if(verbose) { cat("Beginning Sampler \n") }
  
  for(i in 1:N.iter){
    
    ###############################################################
    ## update b-spline parameters for fixed effects
    ###############################################################

    mean.cur = as.vector(t(pcaef.cur))

    sigma = solve( (1/sig2.me) * sig.W + kronecker(diag(1/lambda.bw, length(lambda.bw), length(lambda.bw)), P.mat)  )
    mu = (1/sig2.me) * sigma %*% (t.designmat.W %*%  (Y.vec - mean.cur))
      
    bw = matrix(mvrnorm(1, mu = mu, Sigma = sigma), nrow = Kt, ncol = p)
    beta.cur = t(bw) %*% t(Theta)
    fixef.cur = W.des %*% beta.cur
   
    ###############################################################
    ## update b-spline parameters for PC basis functions
    ###############################################################

    mean.cur = as.vector(t(fixef.cur))

    sigma = solve( (1/sig2.me) * kronecker(t(c.mat) %*% c.mat, t(Theta)%*% Theta) + kronecker(diag(1/lambda.psi), P.mat  ))
    mu = (1/sig2.me) * sigma %*% t(kronecker(c.mat, Theta)) %*%  (Y.vec - mean.cur)

    bpsi = matrix(mvrnorm(1, mu = mu, Sigma = sigma), nrow = Kt, ncol = Kp)
    psi.cur = t(bpsi) %*% t(Theta)
    
    ppT = psi.cur %*% t(psi.cur)

    ###############################################################
    ## scores for each individual
    ###############################################################

    for(c in 1:I){
      sigma = solve( (1/sig2.me)* ppT + diag(1, Kp, Kp)  )
      mu = (1/sig2.me) * sigma %*% psi.cur %*%  (Y[c,] - fixef.cur[c,] )     
      c.mat[c,] = mvrnorm(1, mu = mu, Sigma = sigma)
    }
    
    pcaef.cur = c.mat %*% psi.cur

    ###############################################################
    ## update variance components
    ###############################################################
      
    ## sigma.me
    Y.cur = fixef.cur +  pcaef.cur
    a.post = Asig + I*D/2
    b.post = Bsig + 1/2*crossprod(as.vector(Y - Y.cur))
    sig2.me = 1/rgamma(1, a.post, b.post)

    ## lambda for beta's
    for(term in 1:p){
      a.post = Aw + Kt/2
      b.post = Bw[term] + 1/2*bw[,term] %*% P.mat %*% bw[,term]
      lambda.bw[term] = 1/rgamma(1, a.post, b.post)
    }

    ## lambda for psi's
    for(K in 1:Kp){
      a.post = Apsi + Kt/2
      b.post = Bpsi + 1/2*bpsi[,K] %*% P.mat %*% bpsi[,K]
      lambda.psi[K] = 1/rgamma(1, a.post, b.post)
    }

    ###############################################################
    ## save this iteration's parameters
    ###############################################################

    BW[,,i] = as.matrix(bw)
    BPSI[,,i] = as.matrix(bpsi)
    C[,,i] = as.matrix(c.mat)
      
    SIGMA[i] = sig2.me
    LAMBDA.BW[i,] = lambda.bw
    LAMBDA.PSI[i,] = lambda.psi

    if(i > N.burn){
      y.post[,,i - N.burn] = fixef.cur+pcaef.cur
    }
  
    if(verbose) { if(round(i %% (N.iter/10)) == 0) {cat(".")} }
    
  }
  
  ###############################################################
  ## compute posteriors for this dataset
  ###############################################################

  ## main effects
  beta.pm = beta.LB = beta.UB = matrix(NA, nrow = p, ncol = D)
  for(i in 1:p){
  	beta.post = matrix(NA, nrow = (N.iter - N.burn), ncol = D)
  	for(n in 1:(N.iter - N.burn)){
  	  beta.post[n,] = BW[,i, n + N.burn] %*% t(Theta)
  	}
  	beta.pm[i,] = apply(beta.post, 2, mean)
    beta.LB[i,] = apply(beta.post, 2, quantile, c(.025))
    beta.UB[i,] = apply(beta.post, 2, quantile, c(.975))
  }


  ## FPCA basis functions -- OUT OF DATE
  psi.pm = matrix(NA, nrow = Kp, ncol = D)
  for(i in 1:Kp){
  	psi.post = matrix(NA, nrow = (N.iter - N.burn), ncol = D)
  	for(n in 1:(N.iter - N.burn)){
  	  psi.post[n,] = BPSI[,i, n + N.burn] %*% t(Theta)
  	}
  	psi.pm[i,] = apply(psi.post, 2, mean)
  }

  ## export fitted values
  Yhat = apply(y.post, c(1,2), mean)
  y.LB = apply(y.post, c(1,2), quantile, c(.025))
  y.UB = apply(y.post, c(1,2), quantile, c(.975))
  
  data = if(is.null(data)) { mf_fixed }  else { data }
  
  ret = list(beta.pm, beta.UB, beta.LB, Yhat, mt_fixed, data, psi.pm)
  names(ret) = c("beta.hat", "beta.UB", "beta.LB", "Yhat", "terms", "data", "psi.pm")
  class(ret) = "fosr"
  ret
  
}

```

