---
title: "explore tidyfun"
author: "Gaeun Kim"
date: "1/4/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidyfun)
library(refund)
library(ggplot2)
library(reshape2)
library(mgcv)
source("../function/quadWeights.R")
source("../function/fpca.tfd.R")
```

First, let's read in our tf datasets.
```{r}
temp_tfd = load("../data/temp_tfd.RData")
handw_tfd= load("../data/handw_tfd.RData")

DTI = refund::DTI

dti = with(refund::DTI, 
  data.frame(id = ID, sex = sex, 
    case = factor(ifelse(case, "MS", "control")))) %>% as.tbl %>% 
        mutate(cca = tfd(DTI$cca, seq(0,1, l = 93), signif = 2) %>%
                     tfd(arg = seq(0,1,l = 93)),
               rcst = tfd(DTI$rcst, seq(0, 1, l = 55), signif = 3))
```

fpca.tfd function examples
```{r}
fit.cca = fpca.tfd(data = dti, col = rcst)

fit.mu = data.frame(mu = fit.cca$mu,
                    n = 1:ncol(fit.cca$Yhat))
fit.basis = data.frame(phi = fit.cca$efunctions, #the FPC basis functions.
                       n = 1:ncol(fit.cca$Yhat))

## plot estimated mean function
ggplot(fit.mu, aes(x = n, y = mu)) + geom_path() 

## plot the first two estimated basis functions
fit.basis.m = melt(fit.basis, id = 'n')
ggplot(subset(fit.basis.m, variable %in% c('phi.1', 'phi.2')), aes(x = n,
y = value, group = variable, color = variable)) + geom_path()

```

putting cd4 dataset to fpca.tfd function

```{r}
data(cd4)
# CD4 cell counts for 366 subjects between months -18 and 42 since seroconversion. Each subject's observations are contained in a single row.
# subject * weeks

cd4_tfd = cd4 %>%
  tfd()
cd4_df = data_frame(cd4_tfd, id = 1:366) %>%
  select(cd4 = cd4_tfd, everything())

fit.cd4 = cd4_df %>%
  fpca.tfd(data =., col = cd4)

fit.mu = data.frame(mu = fit.cd4$mu,
                    n = 1:ncol(fit.cd4$Yhat))
fit.basis = data.frame(phi = fit.cd4$efunctions, #the FPC basis functions.
                       n = 1:ncol(fit.cd4$Yhat))

## plot estimated mean function
ggplot(fit.mu, aes(x = n, y = mu)) + geom_path() 
```

cd4 example on the refund page
```{r}
Fit.MM = fpca.sc(cd4, var = TRUE, simul = TRUE)

Fit.mu = data.frame(mu = Fit.MM$mu,
                    d = as.numeric(colnames(cd4)))
Fit.basis = data.frame(phi = Fit.MM$efunctions, # d × npc matrix of estimated eigenfunctions of the functional covariance, i.e., the FPC basis functions.
                       d = as.numeric(colnames(cd4)))

## for one subject, examine curve estimate, pointwise and simultaneous itervals
EX = 1
EX.MM = data.frame(fitted = Fit.MM$Yhat[EX,],
           ptwise.UB = Fit.MM$Yhat[EX,] + 1.96 * sqrt(Fit.MM$diag.var[EX,]),
           ptwise.LB = Fit.MM$Yhat[EX,] - 1.96 * sqrt(Fit.MM$diag.var[EX,]),
           simul.UB = Fit.MM$Yhat[EX,] + Fit.MM$crit.val[EX] * sqrt(Fit.MM$diag.var[EX,]),
           simul.LB = Fit.MM$Yhat[EX,] - Fit.MM$crit.val[EX] * sqrt(Fit.MM$diag.var[EX,]),
           d = as.numeric(colnames(cd4)))

## plot data for one subject, with curve and interval estimates
EX.MM.m = melt(EX.MM, id = 'd')
ggplot(EX.MM.m, aes(x = d, y = value, group = variable, color = variable, linetype = variable)) +
  geom_path() +
  scale_linetype_manual(values = c(fitted = 1, ptwise.UB = 2,
                        ptwise.LB = 2, simul.UB = 3, simul.LB = 3)) +
  scale_color_manual(values = c(fitted = 1, ptwise.UB = 2,
                     ptwise.LB = 2, simul.UB = 3, simul.LB = 3)) +
  labs(x = 'Months since seroconversion', y = 'Total CD4 Cell Count')

## plot estimated mean function
ggplot(Fit.mu, aes(x = d, y = mu)) + geom_path() +
  labs(x = 'Months since seroconversion', y = 'Total CD4 Cell Count')

## plot the first two estimated basis functions
Fit.basis.m = melt(Fit.basis, id = 'd')
ggplot(subset(Fit.basis.m, variable %in% c('phi.1', 'phi.2')), aes(x = d,
y = value, group = variable, color = variable)) + geom_path()

```
looks same as the tfd version!

bayes_fosr:
"Wrapper function that implements several approaches to Bayesian function on scalar regression."

Exammple code from refund

* I am getting errors regarding gibbs so I removed it from the sample for now.
* error message: Error in tx[, ind.cur] %*% siginv : non-conformable arguments

I won't alter DTI like they did in refund because I want it to be same as dti, a tdf format of DTI.
```{r }
##### Cross-sectional real-data examples #####

## organize data
data(DTI)
#DTI = subset(DTI, select = c(cca, case, pasat))
#DTI = DTI[complete.cases(DTI),]
#DTI$gender = factor(sample(c("male","female"), dim(DTI)[1], replace = TRUE))
#DTI$status = factor(sample(c("RRMS", "SPMS", "PPMS"), dim(DTI)[1], replace = TRUE))

## fit models
default = bayes_fosr(cca ~ pasat, data = DTI)
VB = bayes_fosr(cca ~ pasat, data = DTI, Kp = 4, Kt = 10)
#Gibbs = bayes_fosr(cca ~ pasat, data = DTI, Kt = 10, est.method = "Gibbs", cov.method = #"Wishart", N.iter = 500, N.burn = 200)
OLS = bayes_fosr(cca ~ pasat, data = DTI, Kt = 10, est.method = "OLS")
GLS = bayes_fosr(cca ~ pasat, data = DTI, Kt = 10, est.method = "GLS")

## plot results
models = c("default", "VB", "OLS", "GLS")
intercepts = sapply(models, function(u) get(u)$beta.hat[1,])
slopes = sapply(models, function(u) get(u)$beta.hat[2,])

plot.dat = melt(intercepts); colnames(plot.dat) = c("grid", "method", "value")
ggplot(plot.dat, aes(x = grid, y = value, group = method, color = method)) + 
   geom_path() + theme_bw()

plot.dat = melt(slopes); colnames(plot.dat) = c("grid", "method", "value")
ggplot(plot.dat, aes(x = grid, y = value, group = method, color = method)) + 
   geom_path() + theme_bw()

#Gibbs - Error in tx[, ind.cur] %*% siginv : non-conformable arguments
# removed Gibbs
```

ols_cs:
"Fitting function for function-on-scalar regression for cross-sectional data."
```{r original ols}
dti.ols = ols_cs(cca ~ pasat, data = DTI, Kt = 10)
```

Below is a ols function for cross-sectional tfd datasets.
```{r}
source("../function/ols_cs_tfd.R")
```

I will put in our tf dataset that we made in the beginning, dti.

```{r fit tf ols}
dti["pasat"] = DTI$pasat

dti.ols.prac = ols_cs_tfd(cca ~ pasat, col = cca, data = dti, Kt = 10)

models = c("dti.ols", "dti.ols.prac")
intercepts = sapply(models, function(u) get(u)$beta.hat[1,])
slopes = sapply(models, function(u) get(u)$beta.hat[2,])

plot.dat = melt(intercepts); colnames(plot.dat) = c("grid", "method", "value")
ggplot(plot.dat, aes(x = grid, y = value, group = method, color = method)) + 
   geom_path() + theme_bw()

plot.dat = melt(slopes); colnames(plot.dat) = c("grid", "method", "value")
ggplot(plot.dat, aes(x = grid, y = value, group = method, color = method)) + 
   geom_path() + theme_bw()
```

**Issue**

the dataset altered inside ols_cs_tfd seems to be same as DTI but it's giving slightly different results as we can see in the graph above. 

checking if the datasets(altered dti and original DTI) are the same:
```{r}
all(dti %>%
      pull(cca) %>%
      as.data.frame() %>%
      spread(key = arg, value = value) %>%
      select(-id) %>%
      as.matrix() == DTI$cca, na.rm = TRUE)

all(dti$pasat == DTI$pasat, na.rm = TRUE)
```
So the cca from dti altered to a matrix in the function is the same as cca in DTI as well as pasat in dti and in DTI. They also have equal dim size.

```{r}
plot_fit_models = function(models) {
  
  models = models
  intercepts = sapply(models, function(u) get(u)$beta.hat[1,])
  slopes = sapply(models, function(u) get(u)$beta.hat[2,])

  plot.dat = melt(intercepts); colnames(plot.dat) = c("grid", "method", "value")
  p1 = ggplot(plot.dat, aes(x = grid, y = value, group = method, color = method)) + 
   geom_path() + theme_bw() + ggtitle("intercepts")

  plot.dat = melt(slopes); colnames(plot.dat) = c("grid", "method", "value")
  p2 = ggplot(plot.dat, aes(x = grid, y = value, group = method, color = method)) + 
   geom_path() + theme_bw() + ggtitle("slopes")
  
 gridExtra::grid.arrange(p1,p2)
}
```


Just to make sure if cca is the cause of the difference I will put DTI's cca inside dti.
```{r}
dti_same = dti
dti_same["cca"] = DTI["cca"]

dti_same_ols = ols_cs(cca ~ pasat, data = dti_same, Kt = 10)

models = c("dti.ols", "dti_same_ols")
plot_fit_models(models)
```

So, cca is the cause of the difference.

Let's see what we get if we change tdf dti to the original version and fit to cs.ols function.
```{r}
dti_testing = dti
dti_testing["cca"] = dti$cca %>% 
  as.data.frame() %>%
  spread(key = arg, value = value) %>%
  select(-id) %>%
  as.matrix() 

dti.testing.ols = ols_cs(cca ~ pasat, data = dti_testing, Kt = 10)
```

```{r}
dti_cca = dti_testing$cca %>% as.data.frame()
DTI_cca = DTI$cca %>% as.data.frame()
rownames(dti_cca) = rownames(DTI_cca)
colnames(dti_cca) = colnames(DTI_cca)

dti_cca %>% is.na() %>% sum() #0 
DTI_cca %>% is.na() %>% sum() #36
```

So I see that while both cca columns have the sme values for non-missing data, cca at dti have no missing values while DTI has 36 missings.

* Question: does tfd function remove missings?

```{r}
which(is.na(DTI_cca), arr.ind=TRUE)
dti_cca[321, 45]
DTI_cca[321, 45]
```

Looking at one of the missings in DTI, I think the tfd function impute/approximate the missings.

gibbs_cs_fpca:
"Cross-sectional FoSR using a Gibbs sampler and FPCA"
```{r}
#gibbs_cs_fpca(cca ~ pasat, data = DTI, Kt = 10)
#gibbs_cs_fpca_prac(cca ~ pasat, data = DTI, Kt = 10)

#keeps on running...
```

```{r eval = FALSE, echo = FALSE}
library(MASS)
library(lme4)

# gibbs_cs_fpca_prac = 
```


