---
title: "explore tidyfun"
author: "Sara Kim"
date: "1/4/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidyfun)
library(refund)
library(ggplot2)
library(reshape2)
library(mgcv)
source("../function/quadWeights.R")
source("../function/fpca.tfd.R")
```

First, let's read in our tf datasets.
```{r}
temp_tfd = load("../data/temp_tfd.RData")
handw_tfd= load("../data/handw_tfd.RData")

DTI = refund::DTI

dti = with(refund::DTI, 
  data.frame(id = ID, sex = sex, 
    case = factor(ifelse(case, "MS", "control")))) %>% as.tbl %>% 
        mutate(cca = tfd(DTI$cca, seq(0,1, l = 93), signif = 2) %>%
                     tfd(arg = seq(0,1,l = 93)),
               rcst = tfd(DTI$rcst, seq(0, 1, l = 55), signif = 3))
```

fpca.tfd function examples
```{r}
fit.cca = fpca.tfd(data = dti, col = rcst)

fit.mu = data.frame(mu = fit.cca$mu,
                    n = 1:ncol(fit.cca$Yhat))
fit.basis = data.frame(phi = fit.cca$efunctions, #the FPC basis functions.
                       n = 1:ncol(fit.cca$Yhat))

## plot estimated mean function
ggplot(fit.mu, aes(x = n, y = mu)) + geom_path() 

## plot the first two estimated basis functions
fit.basis.m = melt(fit.basis, id = 'n')
ggplot(subset(fit.basis.m, variable %in% c('phi.1', 'phi.2')), aes(x = n,
y = value, group = variable, color = variable)) + geom_path()

```

putting cd4 dataset to fpca.tfd function

```{r}
data(cd4)
# CD4 cell counts for 366 subjects between months -18 and 42 since seroconversion. Each subject's observations are contained in a single row.
# subject * weeks

cd4_tfd = cd4 %>%
  tfd()
cd4_df = data_frame(cd4_tfd, id = 1:366) %>%
  select(cd4 = cd4_tfd, everything())

fit.cd4 = cd4_df %>%
  fpca.tfd(data =., col = cd4)

fit.mu = data.frame(mu = fit.cd4$mu,
                    n = 1:ncol(fit.cd4$Yhat))
fit.basis = data.frame(phi = fit.cd4$efunctions, #the FPC basis functions.
                       n = 1:ncol(fit.cd4$Yhat))

## plot estimated mean function
ggplot(fit.mu, aes(x = n, y = mu)) + geom_path() 
```

cd4 example on the refund page
```{r}
Fit.MM = fpca.sc(cd4, var = TRUE, simul = TRUE)

Fit.mu = data.frame(mu = Fit.MM$mu,
                    d = as.numeric(colnames(cd4)))
Fit.basis = data.frame(phi = Fit.MM$efunctions, # d × npc matrix of estimated eigenfunctions of the functional covariance, i.e., the FPC basis functions.
                       d = as.numeric(colnames(cd4)))

## for one subject, examine curve estimate, pointwise and simultaneous itervals
EX = 1
EX.MM = data.frame(fitted = Fit.MM$Yhat[EX,],
           ptwise.UB = Fit.MM$Yhat[EX,] + 1.96 * sqrt(Fit.MM$diag.var[EX,]),
           ptwise.LB = Fit.MM$Yhat[EX,] - 1.96 * sqrt(Fit.MM$diag.var[EX,]),
           simul.UB = Fit.MM$Yhat[EX,] + Fit.MM$crit.val[EX] * sqrt(Fit.MM$diag.var[EX,]),
           simul.LB = Fit.MM$Yhat[EX,] - Fit.MM$crit.val[EX] * sqrt(Fit.MM$diag.var[EX,]),
           d = as.numeric(colnames(cd4)))

## plot data for one subject, with curve and interval estimates
EX.MM.m = melt(EX.MM, id = 'd')
ggplot(EX.MM.m, aes(x = d, y = value, group = variable, color = variable, linetype = variable)) +
  geom_path() +
  scale_linetype_manual(values = c(fitted = 1, ptwise.UB = 2,
                        ptwise.LB = 2, simul.UB = 3, simul.LB = 3)) +
  scale_color_manual(values = c(fitted = 1, ptwise.UB = 2,
                     ptwise.LB = 2, simul.UB = 3, simul.LB = 3)) +
  labs(x = 'Months since seroconversion', y = 'Total CD4 Cell Count')

## plot estimated mean function
ggplot(Fit.mu, aes(x = d, y = mu)) + geom_path() +
  labs(x = 'Months since seroconversion', y = 'Total CD4 Cell Count')

## plot the first two estimated basis functions
Fit.basis.m = melt(Fit.basis, id = 'd')
ggplot(subset(Fit.basis.m, variable %in% c('phi.1', 'phi.2')), aes(x = d,
y = value, group = variable, color = variable)) + geom_path()

```
looks same as the tfd version!

Sample Code from Refund
bayes_fosr:
"Wrapper function that implements several approaches to Bayesian function on scalar regression."

I won't alter DTI like they did in refund because I want it to be same as dti, a tdf format of DTI.
```{r }
##### Cross-sectional real-data examples #####

## organize data
data(DTI)
#DTI = subset(DTI, select = c(cca, case, pasat))
#DTI = DTI[complete.cases(DTI),]
#DTI$gender = factor(sample(c("male","female"), dim(DTI)[1], replace = TRUE))
#DTI$status = factor(sample(c("RRMS", "SPMS", "PPMS"), dim(DTI)[1], replace = TRUE))

## fit models
default = bayes_fosr(cca ~ pasat, data = DTI)
VB = bayes_fosr(cca ~ pasat, data = DTI, Kp = 4, Kt = 10)
# to run Gibbs download refund from devtools
Gibbs = bayes_fosr(cca ~ pasat, data = DTI, Kt = 10, est.method = "Gibbs", cov.method = "FPCA", N.iter = 500, N.burn = 200)
## Gibbs Wishart
Gibbs_wish = bayes_fosr(cca ~ pasat, data = DTI, Kt = 10, est.method = "Gibbs", cov.method = "Wishart", N.iter = 500, N.burn = 200)
OLS = bayes_fosr(cca ~ pasat, data = DTI, Kt = 10, est.method = "OLS")
GLS = bayes_fosr(cca ~ pasat, data = DTI, Kt = 10, est.method = "GLS")

## plot results
models = c("default", "VB", "Gibbs", "OLS", "GLS")
intercepts = sapply(models, function(u) get(u)$beta.hat[1,])
slopes = sapply(models, function(u) get(u)$beta.hat[2,])

plot.dat = melt(intercepts); colnames(plot.dat) = c("grid", "method", "value")
ggplot(plot.dat, aes(x = grid, y = value, group = method, color = method)) + 
   geom_path() + theme_bw()

plot.dat = melt(slopes); colnames(plot.dat) = c("grid", "method", "value")
ggplot(plot.dat, aes(x = grid, y = value, group = method, color = method)) + 
   geom_path() + theme_bw()
```

ols_cs:
"Fitting function for function-on-scalar regression for cross-sectional data."
```{r original ols}
dti.ols = ols_cs(cca ~ pasat, data = DTI, Kt = 10)
```

Below is a ols function for cross-sectional tfd datasets.
```{r}
source("../function/ols_cs_tfd.R")
```

I will put in our tf dataset that we made in the beginning, dti.

```{r fit tf ols}
dti["pasat"] = DTI$pasat

dti.ols.prac = ols_cs_tfd(cca ~ pasat, data = dti, Kt = 10)

models = c("dti.ols", "dti.ols.prac")
intercepts = sapply(models, function(u) get(u)$beta.hat[1,])
slopes = sapply(models, function(u) get(u)$beta.hat[2,])

plot.dat = melt(intercepts); colnames(plot.dat) = c("grid", "method", "value")
ggplot(plot.dat, aes(x = grid, y = value, group = method, color = method)) + 
   geom_path() + theme_bw()

plot.dat = melt(slopes); colnames(plot.dat) = c("grid", "method", "value")
ggplot(plot.dat, aes(x = grid, y = value, group = method, color = method)) + 
   geom_path() + theme_bw()
```

The dataset altered inside ols_cs_tfd seems to be same as DTI except the null values inside DTI have been imputed in the tfd. I belive because of that It's giving slightly different results as we can see in the graph above. 

```{r}
dti_testing = dti
dti_testing["cca"] = dti$cca %>% 
  as.data.frame() %>%
  spread(key = arg, value = value) %>%
  select(-id) %>%
  as.matrix() 

dti_cca = dti_testing$cca %>% as.data.frame()
DTI_cca = DTI$cca %>% as.data.frame()
rownames(dti_cca) = rownames(DTI_cca)
colnames(dti_cca) = colnames(DTI_cca)

dti_cca %>% is.na() %>% sum() #0 
DTI_cca %>% is.na() %>% sum() #36
```

So I see that while both cca columns have the sme values for non-missing data, cca at dti have no missing values while DTI has 36 missings.

gibbs_cs_fpca:
"Cross-sectional FoSR using a Gibbs sampler and FPCA"
```{r}
Gibbs_bayes = bayes_fosr(cca ~ pasat, data = DTI, Kt = 10, est.method = "Gibbs", cov.method = "FPCA", N.iter = 500, N.burn = 200)

gibbs = gibbs_cs_fpca(cca ~ pasat, data = DTI, Kt = 10, N.iter = 500, N.burn = 200)
```

```{r}
source("../function/gibbs_cs_fpca_tfd.R")
```

```{r}
gibbs_dti = gibbs_cs_fpca_tfd(cca ~ pasat, data = dti, Kt = 10, N.iter = 500, N.burn = 200)
```

```{r}
source("../function/gibbs_cs_wish_tfd.R")
```

```{r}
gibbs_dti_wish = gibbs_cs_wish_tfd(cca ~ pasat, data = dti, Kt = 10, N.iter = 500, N.burn = 200)
```

```{r}
models = c("Gibbs_bayes", "gibbs", "gibbs_dti", "gibbs_dti_wish")
intercepts = sapply(models, function(u) get(u)$beta.hat[1,])
slopes = sapply(models, function(u) get(u)$beta.hat[2,])

plot.dat = melt(intercepts); colnames(plot.dat) = c("grid", "method", "value")
ggplot(plot.dat, aes(x = grid, y = value, group = method, color = method)) + 
   geom_path() + theme_bw()

plot.dat = melt(slopes); colnames(plot.dat) = c("grid", "method", "value")
ggplot(plot.dat, aes(x = grid, y = value, group = method, color = method)) + 
   geom_path() + theme_bw()
```

So I've graphed intercepts and slopes from four functions using Gibbs sampler. They all seems to differ slightly. Strangely, Gibbs_bayes and gibbs were fitted using the same `gibbs_cs_fpca` function but still gives different results. (Why?)